# Note to Self: There is no Rcpp const!

```{r load, echo = FALSE, eval = FALSE}
devtools::load_all (".", export_all = FALSE)
devtools::document (".")
```

`Rcpp` passes all internals objects by reference only, never by value. This repo
demonstrates the implications of this as a clear and simple Note To Self. It
relies on these `C++` functions
```
// [[Rcpp::export]]
void rcpp_test1(const Rcpp::DataFrame df) {
    Rcpp::IntegerVector x = df ["x"];
    x = x + 1;
}

// [[Rcpp::export]]
void rcpp_test2(const Rcpp::DataFrame df) {
    std::vector <int> x = df ["x"];
    for (auto i: x)
        i++;
}
```

```{r, message = FALSE}
devtools::load_all (".", export_all = FALSE)
df <- data.frame (x = 1:5, y = 1:5)
test1 (df)
df
```
Running `test1()` alters the values of `df` because the internal
`Rcpp::IntegerVector` object is constructed strictly by reference only.
Crucially, in doing so, `Rcpp` also **complete ignores the `const` directive**.
In contrast, the `rcpp_test2` function implements an implicit copy-by-value
through the typecast to `std::vector`, and so
```{r}
test2 (df)
df
```
An alternative is `Rcpp::clone()`, as demonstrated in the third function:
```
// [[Rcpp::export]]
void rcpp_test3(const Rcpp::DataFrame df) {
    const Rcpp::IntegerVector x = df ["x"];
    Rcpp::IntegerVector x2 = Rcpp::clone (x);
    x2 = x2 + 1;
}
```
This also leaves the original unmodified:
```{r}
test3 (df)
df
```

## moral of the story

The following code is pointless:
```
void rcpp_stuff(const Rcpp::<object> obj) { ... }
```
because `const` will just be ignored anyway. Never forget that! Easy way is
never to pretend that `const Rcpp::<object>` has any meaning, and just write
this instead:
```
void rcpp_stuff(Rcpp::<object> obj) { ... }
```
